---
title: "LiDAR and Sentinel-2 LAI Random Forest Evolution Analysis"
author: "Nathan Corroyez"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  pdf_document: default
---

## Introduction

This notebook performs Random Forest analysis on LiDAR and Sentinel-2 LAI data for different forest sites using various metrics. We explore different cases by including/excluding dummy variables and analyzing two target variables: `LiDAR_LAI` and `deltaLAI`.

## Pre-processing Parameters

```{r setup, echo=TRUE}
rm(list=ls(all=TRUE)) # Clear the global environment (remove all objects)
gc() # Trigger the garbage collector to free up memory

# Load required libraries
library(terra)
library(ranger)
library(dplyr)
library(ggplot2)

# Pre-processing parameters
sites <- c("Aigoual", "Blois", "Mormal")
# sites <- c("Blois")
# sites <- c("Mormal", "Blois")
results_dir <- "../../03_RESULTS"
output_dir <- "../../04_FIGURES/RF"
forest_composition <- "Not_Masked"  # Can be "Full_Composition", "Deciduous_Only", "Not_Masked"
set.seed(42)  # Ensure reproducibility

```

## Function to load metrics for a given site

```{r, echo=F}
load_metrics_for_site <- function(site, results_dir, forest_composition) {
  # Define the metrics directory for the given site
  metrics_dir <- file.path(results_dir, site, "Metrics", forest_composition)
  
  # Define the names of the metrics
  metric_files <- c("lidarlai_res_10_m.tif", 
                    "s2lai_res_10_m.tif",
                    "lcv_res_10_m.tif",
                    "mean_res_10_m.tif",
                    "max_res_10_m.tif",
                    "lskew_res_10_m.tif",
                    "shade_res_10_m.tif",
                    "slope_res_10_m.tif",
                    "aspect_res_10_m.tif",
                    "cv_res_10_m.tif",
                    "rumple_res_10_m.tif",
                    "vci_res_10_m.tif",
                    "vdr_res_10_m.tif",
                    "gap_fraction_res_10_m.tif",
                    # "ndvi_summer_res_10_m.tif",
                    # "stand_type_raster.tif",
                    # "dtm.tif",
                    "cv_lad_res_10_m.tif"
  )
  
  # Function to load and extract raster values
  load_metric <- function(filename) {
    rast <- terra::rast(file.path(metrics_dir, filename))
    # print(rast)
    values(rast)
  }
  
  # Load all the metrics in a list
  metrics_values <- lapply(metric_files, load_metric)
  
  # Assign meaningful names to the metrics
  names(metrics_values) <- c("lidar_lai", 
                             "s2_lai",
                             "lcv",
                             "mean",
                             "max",
                             "lskew",
                             "shade",
                             "slope",
                             "aspect",
                             "cv",
                             "rumple",
                             "vci",
                             "vdr",
                             "gap_fraction",
                             # "ndvi",
                             # "stands",
                             # "dtm",
                             "cv_lad"
  )
  # print(str(metrics_values))
  
  # Generate names for the PAD profiles
  # pad_files <- sprintf("PAD_Profiles_updated/PAD_%.1f_40.tif", seq(2.5, 39.5, by = 1))

  # Load all the PAD profile metrics in a list
  # pad_metrics_values <- lapply(pad_files, load_metric)

  # Assign meaningful names to the PAD profile metrics
  # names(pad_metrics_values) <- sprintf("PAD_%.1f", seq(2.5, 39.5, by = 1))

  # Combine the primary and PAD metrics
  # metrics_values <- c(metrics_values, pad_metrics_values)
  
  # Extract coordinates (assuming all rasters have the same structure)
  lidar_lai <- terra::rast(file.path(metrics_dir, "lidarlai_res_10_m.tif"))
  coordinates <- xyFromCell(lidar_lai, 1:ncell(lidar_lai))
  
  # Add longitude and latitude to the list of metrics
  metrics_values$lon <- coordinates[, 1]
  metrics_values$lat <- coordinates[, 2]
  
  # Filter out NA values across all metrics
  valid_idx <- complete.cases(do.call(cbind, metrics_values))
  
  # Filter metrics to retain only valid (non-NA) rows
  valid_metrics <- lapply(metrics_values, function(x) x[valid_idx])
  
  # Combine all predictor variables into a data frame
  predictors <- as.data.frame(valid_metrics)
  # predictors$site <- site
  
  return(predictors)
}
```

## Test

```{r}
# Initialize an empty list to store the data frames
all_predictors <- list()

# Load data for all sites
for (site in sites) {
  site_data <- load_metrics_for_site(site, results_dir, forest_composition)
  
  all_predictors[[site]] <- site_data  # Store the data frame in the list
}
# Combine all the data frames into one
combined_predictors <- do.call(rbind, all_predictors)
```

```{r}
partialPlot <- function(obj, pred.data, xname, n.pt = 9, discrete.x = FALSE, 
                        subsample = pmin(1, n.pt * 100 / nrow(pred.data)), which.class = NULL,
                        xlab = deparse(substitute(xname)), ylab = "", type = if (discrete.x) "p" else "b",
                        main = "", rug = TRUE, seed = NULL, show_plot = TRUE, target_var,
                        save_path = NULL, ...) {
  stopifnot(dim(pred.data) >= 1)
  
  # Subsample the data if needed
  if (subsample < 1) {
    if (!is.null(seed)) {
      set.seed(seed)
    } 
    n <- nrow(pred.data)
    picked <- sample(n, trunc(subsample * n))
    pred.data <- pred.data[picked, , drop = FALSE]
  }
  
  xv <- pred.data[, xname]
  
  # Define quantiles from 0.1 to 0.9 (9 points)
  if (discrete.x) {
    x <- unique(xv)
  } else {
    x <- quantile(xv, probs = seq(0.1, 0.9, length.out = n.pt), names = FALSE)  # Quantiles from 0.1 to 0.9
  }
  
  y <- numeric(length(x))
  
  isRanger <- inherits(obj, "ranger")
  isLm <- inherits(obj, "lm") | inherits(obj, "lmrob") | inherits(obj, "lmerMod")
  
  # Calculate partial dependence by predicting for each x[i]
  for (i in seq_along(x)) {
    pred.data[, xname] <- x[i]

    if (isRanger) {
      if (!is.null(which.class)) {
        if (obj$treetype != "Probability estimation") {
          stop("Choose probability = TRUE when fitting ranger multiclass model") 
        }
        preds <- predict(obj, pred.data)$predictions[, which.class]
      }
      else {
        preds <- predict(obj, pred.data)$predictions
      }
    } else if (isLm) {
      preds <- predict(obj, pred.data) 
    } else {
      if (!is.null(which.class)) {
        preds <- predict(obj, pred.data, reshape = TRUE)[, which.class + 1] 
      } else {
        preds <- predict(obj, pred.data)
      }
    }
    
    y[i] <- mean(preds)
  }
  
  if (show_plot) {
    # Save the plot if save_path is provided
    if (!is.null(save_path)) {
      png(filename = file.path(save_path, paste0("pp_", xname, "_target_", target_var, ".png")), width = 1920, height = 1080, res = 300)  
    }

    # Create the plot
    plot(x, y, xlab = target_var, ylab = ylab, 
         main = paste("Partial Plot for", xname, "target var", target_var), 
         type = type, ...)
    
    if (!is.null(save_path)) {
      dev.off()  # Close the graphical device
    }
  }
  
  # Return the partial dependence data for further use
  data.frame(x = x, y = y)
}
```

```{r}
biPartialPlot <- function(obj, pred.data, xname1, xname2, n.pt = 9, discrete.x1 = FALSE, discrete.x2 = FALSE, 
                          subsample = pmin(1, n.pt * 100 / nrow(pred.data)), which.class = NULL,
                          xlab1 = deparse(substitute(xname1)), xlab2 = deparse(substitute(xname2)), 
                          ylab = "", main = "", rug = TRUE, seed = NULL, show_plot = TRUE, target_var,
                          save_path = NULL, ...) {
  
  stopifnot(dim(pred.data) >= 1)
  
  # Subsample the data if needed
  if (subsample < 1) {
    if (!is.null(seed)) {
      set.seed(seed)
    } 
    n <- nrow(pred.data)
    picked <- sample(n, trunc(subsample * n))
    pred.data <- pred.data[picked, , drop = FALSE]
  }
  
  xv1 <- pred.data[, xname1]
  xv2 <- pred.data[, xname2]
  
  # Define quantiles for each predictor
  if (discrete.x1) {
    x1 <- unique(xv1)
  } else {
    x1 <- quantile(xv1, probs = seq(0.1, 0.9, length.out = n.pt), names = FALSE)
  }
  
  if (discrete.x2) {
    x2 <- unique(xv2)
  } else {
    x2 <- quantile(xv2, probs = seq(0.1, 0.9, length.out = n.pt), names = FALSE)
  }
  
  # Create a grid for x1 and x2
  grid <- expand.grid(x1 = x1, x2 = x2)
  y <- numeric(nrow(grid))
  
  isRanger <- inherits(obj, "ranger")
  isLm <- inherits(obj, "lm") | inherits(obj, "lmrob") | inherits(obj, "lmerMod")
  
  # Calculate partial dependence by predicting for each combination of x1 and x2
  for (i in seq_len(nrow(grid))) {
    pred.data[, xname1] <- grid$x1[i]
    pred.data[, xname2] <- grid$x2[i]

    if (isRanger) {
      if (!is.null(which.class)) {
        if (obj$treetype != "Probability estimation") {
          stop("Choose probability = TRUE when fitting ranger multiclass model") 
        }
        preds <- predict(obj, pred.data)$predictions[, which.class]
      }
      else {
        preds <- predict(obj, pred.data)$predictions
      }
    } else if (isLm) {
      preds <- predict(obj, pred.data) 
    } else {
      if (!is.null(which.class)) {
        preds <- predict(obj, pred.data, reshape = TRUE)[, which.class + 1] 
      } else {
        preds <- predict(obj, pred.data)
      }
    }
    
    y[i] <- mean(preds)
  }
  
  # Plotting the results
  if (show_plot) {
    if (!is.null(save_path)) {
      png(filename = file.path(save_path, paste0("bi_partial_", xname1, "_", xname2, "_target_", target_var, ".png")), width = 1920, height = 1080, res = 300)  
    }

    # Create the plot
    filled.contour(x1, x2, matrix(y, nrow = length(x1)), 
                   xlab = xlab1, ylab = xlab2, 
                   main = paste("Bi-Partial Plot for", xname1, "and", xname2,
                                "target var", target_var), ...)
    
    if (!is.null(save_path)) {
      dev.off()  # Close the graphical device
    }
  }
  
  # Return the partial dependence data for further use
  data.frame(x1 = rep(x1, each = length(x2)), x2 = rep(x2, times = length(x1)), y = y)
}
```


## Function to run different RF scenarios

```{r}
run_rf_cases <- function(sites, results_dir, output_dir, 
                         forest_composition, use_dummies, target_var,
                         subsampling = FALSE, save_results = TRUE) {
  
  # Initialize an empty list to store the data frames
  all_predictors <- list()
  
  # # Load data for all sites
  for (site in sites) {
    site_data <- load_metrics_for_site(site, results_dir, forest_composition)  # No dummies here
    if (target_var == "deltaLAI") {
      site_data$deltaLAI <- site_data$lidar_lai - site_data$s2_lai
      site_data$lidar_lai <- NULL
      # site_data$s2_lai <- NULL
    }

    all_predictors[[site]] <- site_data  # Store the data frame in the list
  }
  # combined_predictors <- load_metrics_for_site(sites, results_dir, forest_composition)
  # Combine all the data frames into one
  combined_predictors <- do.call(rbind, all_predictors)
  
  # If using dummies, create dummy variables for the 'site' variable
  if (use_dummies) {
    # Convert the 'site' and 'stands' columns to factors
    combined_predictors$site <- as.factor(combined_predictors$site)
    combined_predictors$stands <- as.factor(combined_predictors$stands)
    # Create dummy variables for 'site' and remove the original 'site' column
    dummy_vars_site <- model.matrix(~ site - 1, data = combined_predictors)
    colnames(dummy_vars_site) <- gsub("site", "site_", colnames(dummy_vars_site))
    
    # Create dummy variables for 'stands' and remove the original 'stands' column
    dummy_vars_stands <- model.matrix(~ stands - 1, data = combined_predictors)
    colnames(dummy_vars_stands) <- gsub("stands", "stands_", colnames(dummy_vars_stands))
    
    # Combine the dummy variables with the rest of the predictors
    combined_predictors <- cbind(combined_predictors[, !(colnames(combined_predictors) %in% c(
      "site",
      "stands"))],
                                  as.data.frame(dummy_vars_site),
                                  as.data.frame(dummy_vars_stands))
  }
  
  # Shuffle
  if (subsampling){
    combined_predictors <- combined_predictors[sample(nrow(combined_predictors), 1000), ]
  }
  else {
    combined_predictors <- combined_predictors[sample(nrow(combined_predictors)), ]
  }
  # print(str(combined_predictors))
  
  # Select the appropriate target variable for the RF formula
  if (target_var == "lidar_lai") {
    formula <- lidar_lai ~ .
  } else if (target_var == "deltaLAI") {
    formula <- deltaLAI ~ .
  } else if (target_var == "s2_lai") {
    formula <- s2_lai ~ .
  }
  
  # Fit the Random Forest model using ranger
  rf_model <- ranger(
    formula = formula,
    data = combined_predictors,
    num.trees = 300,
    mtry = floor((ncol(combined_predictors) - 1) / 3),
    importance = "permutation",
    min.node.size = 5,
    num.threads = 12,
    splitrule = "extratrees",
    num.random.splits = 10,
    sample.fraction = 0.05
  )
  # print(str(rf_model))
  # Examine a particular tree
  # library(rpart)
  # t <- rpart(formula,combined_predictors)
  # library(rpart.plot)
  # rpart.plot(t)
  # Extract model evaluation metrics
  oob_mse <- round(rf_model$prediction.error, 2)
  r_squared <- round(rf_model$r.squared, 2)
  
  # Extract variable importance
  varImp <- sort(rf_model$variable.importance, decreasing = TRUE)

  # Plot and save the variable importance plot
  if (save_results) {
    plot_file <- file.path(output_dir,
                           paste0(target_var,
                                  ifelse(use_dummies,
                                         "_with_dummies", "_no_dummies"),
                                  "_importance.png"))
    png(plot_file, width = 1920, height = 1080, res = 300)
  }

  barplot(varImp, main = paste("Variable Importance for",
                               target_var,
                               ifelse(use_dummies,
                                      "with Dummies", "without Dummies")),
          las = 2, cex.names = 0.7)


  # Add MSE and R² to the plot as text
  # mse_text <- paste("OOB MSE:", oob_mse)
  # r2_text <- paste("R²:", r_squared)

  # Add the text to the plot
  # text(x = 10, y = max(varImp) - 1,
  #      labels = paste(mse_text, "\n", r2_text),
  #      pos = 4, cex = 1.6, col = "blue")


  if (save_results) {
    dev.off()  # Close the device if saving
  }
  
  # Save the model's key metrics to a file
  if (save_results) {
    metrics_file <- file.path(output_dir, 
                              paste0(target_var, 
                                     ifelse(use_dummies, 
                                            "_with_dummies", "_no_dummies"), 
                                     "_metrics.txt"))
    writeLines(c(
      paste("OOB MSE:", oob_mse),
      paste("R Squared:", r_squared)
    ), con = metrics_file)
  } else {
    # Print the results to the console if not saving
    cat("OOB MSE:", oob_mse, "\n")
    cat("R Squared:", r_squared, "\n")
  }
  for (var in colnames(combined_predictors[, setdiff(names(combined_predictors), target_var)])){
    partialPlot(obj = rf_model, 
              pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
              xname = var, 
              save_path = output_dir,
              target_var = target_var,
              show_plot = TRUE)
  }
  # partialPlot(obj = rf_model, 
  #             pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
  #             xname = "lcv", 
  #             save_path = output_dir,
  #             target_var = target_var,
  #             show_plot = TRUE)
  # partialPlot(obj = rf_model, 
  #             pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
  #             xname = "mean", 
  #             save_path = output_dir,
  #             target_var = target_var,
  #             show_plot = TRUE)
  # partialPlot(obj = rf_model, 
  #             pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
  #             xname = "gap_fraction", 
  #             save_path = output_dir,
  #             target_var = target_var,
  #             show_plot = TRUE)
  # partialPlot(obj = rf_model, 
  #             pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
  #             xname = "max", 
  #             save_path = output_dir,
  #             target_var = target_var,
  #             show_plot = TRUE)
  # partialPlot(obj = rf_model, 
  #             pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
  #             xname = "shade", 
  #             save_path = output_dir,
  #             target_var = target_var,
  #             show_plot = TRUE)
  biPartialPlot(obj = rf_model, 
                pred.data = combined_predictors[, setdiff(names(combined_predictors), target_var)], 
                xname1 = "lcv", 
                xname2 = "mean", 
                save_path = output_dir,
                target_var = target_var, 
                show_plot = TRUE)
  
  # Remove the model from memory to free up space
  rm(rf_model)
  gc()  # Force garbage collection
  
  return(list(oob_mse = oob_mse,
              var_explained = r_squared,
              variable_importance = varImp
              )
         )
}
```

```{r}
run_quantile_rf_cases <- function(sites, results_dir, output_dir, 
                         forest_composition, use_dummies, target_var,
                         subsampling = FALSE, save_results = TRUE) {
  
  # Initialize an empty list to store the data frames
  all_predictors <- list()
  
  # Load data for all sites
  for (site in sites) {
    site_data <- load_metrics_for_site(site, results_dir, forest_composition)
    if (target_var == "deltaLAI") {
      site_data$deltaLAI <- site_data$lidar_lai - site_data$s2_lai
      site_data$lidar_lai <- NULL
      site_data$s2_lai <- NULL
    }
    
    all_predictors[[site]] <- site_data  # Store the data frame in the list
  }
  
  # Combine all the data frames into one
  combined_predictors <- do.call(rbind, all_predictors)
  
  # If using dummies, create dummy variables for the 'site' variable
  if (use_dummies) {
    combined_predictors$site <- as.factor(combined_predictors$site)
    dummy_vars <- model.matrix(~ site - 1, data = combined_predictors)
    colnames(dummy_vars) <- gsub("site", "site_", colnames(dummy_vars))
    combined_predictors <- cbind(combined_predictors[, !colnames(combined_predictors) %in% "site"], as.data.frame(dummy_vars))
  }
  
  # Shuffle
  if (subsampling){
    combined_predictors <- combined_predictors[sample(nrow(combined_predictors), 10000), ]
  } else {
    combined_predictors <- combined_predictors[sample(nrow(combined_predictors)), ]
  }
  
  # Save the 'site' column separately to associate predictions with the site later
  site_column <- combined_predictors$site
  str(combined_predictors)
  
  # Select the appropriate target variable for the RF formula
  if (target_var == "LiDAR_LAI") {
    formula <- lidar_lai ~ .
  } else if (target_var == "deltaLAI") {
    formula <- deltaLAI ~ .
  }
  
  # Fit the Quantile Forest model using ranger with quantile regression
  rf_model <- ranger(
    formula = formula,
    data = combined_predictors,
    num.trees = 100,
    mtry = floor(ncol(combined_predictors) - 1 / 3),
    importance = "none",
    min.node.size = 5,
    num.threads = 12,
    splitrule = "extratrees",
    num.random.splits = 10,
    sample.fraction = 0.3,
    quantreg = TRUE,
    keep.inbag = TRUE  # Retain inbag information for OOB predictions
  )
  
  # Extract model evaluation metrics
  oob_mse <- round(rf_model$prediction.error, 2)
  r_squared <- round(rf_model$r.squared, 2)
  
  cat("OOB MSE:", oob_mse, "\n")
  cat("R Squared:", r_squared, "\n")
  
  # print(str(rf_model))
  # Predict quantiles (median and IQR) for OOB data
  # predictions <- predict(rf_model, type = "response")
  oob_mse <- round(rf_model$prediction.error, 2)
  predictions_iqr <- predict(rf_model, type = "quantiles", 
                             quantiles = c(0.25, 0.5, 0.75))
  
  # Extract predicted values and confidence intervals
  median_predictions <- predictions_iqr$predictions[,2]
  lower_bound <- predictions_iqr$predictions[,1]  # 25th percentile
  upper_bound <- predictions_iqr$predictions[,3]  # 75th percentile
  ci_width <- upper_bound - lower_bound  # Confidence interval width
  
  # Create a data frame to store predictions and associated site info
  confidence_map <- data.frame(
    lon = combined_predictors$lon,
    lat = combined_predictors$lat,
    median_lai = median_predictions,
    lower_bound = lower_bound,
    upper_bound = upper_bound,
    ci_width = ci_width,
    site = site_column  # Add the 'site' information back for association
  )
  
  # Optional: Save results to the output directory
  if (save_results) {
    write.csv(confidence_map, file = file.path(output_dir, paste0(target_var, "_confidence_map.csv")), row.names = FALSE)
  }
  
  return(list(median_predictions = median_predictions,
              lower_bound = lower_bound,
              upper_bound = upper_bound,
              confidence_map = confidence_map))
}


```

```{r}
# Call the run_rf_cases function
results <- run_quantile_rf_cases(sites = sites, 
                                 results_dir = results_dir, 
                                 output_dir = output_dir,
                                 forest_composition = forest_composition, 
                                 use_dummies = FALSE, 
                                 target_var = "LiDAR_LAI",
                                 subsampling = T, 
                                 save_results = FALSE)

# Inspect the results
head(results$confidence_map)  # View the first few rows of the confidence map
  
# Filter predictions for site1
for (site in sites) {
  # Filter the confidence map for the current site
  site_map <- results$confidence_map %>%
              filter(site == !!site)
  
  # Create output file names for saving
  median_lai_plot_file <- file.path(output_dir, paste0("Median_LAI_", site, ".png"))
  ci_width_plot_file <- file.path(output_dir, paste0("CI_Width_LAI_", site, ".png"))
  
  # Plot the median LAI predictions for the current site
  p1 <- ggplot(site_map, aes(x = lon, y = lat, fill = median_lai)) +
    geom_raster() +
    scale_fill_viridis_c() +
    theme_bw() +
    labs(title = paste("Median LAI Predictions for", site), x = "Longitude", y = "Latitude")
  
  # Save the median LAI plot
  ggsave(median_lai_plot_file, plot = p1, width = 10, height = 8, dpi = 300)
  
  # Plot the confidence interval width for the current site
  p2 <- ggplot(site_map, aes(x = lon, y = lat, fill = ci_width)) +
    geom_raster() +
    scale_fill_viridis_c() +
    theme_bw() +
    labs(title = paste("LAI Confidence Interval Width for", site), x = "Longitude", y = "Latitude")
  
  # Save the confidence interval width plot
  ggsave(ci_width_plot_file, plot = p2, width = 10, height = 8, dpi = 300)
}
```

```{r}
case1 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "s2_lai",
                      subsampling = F, save_results = TRUE)
case2 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "lidar_lai",
                      subsampling = F, save_results = TRUE)
case3 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "deltaLAI",
                      subsampling = F, save_results = F)
```


## 1. LiDAR LAI with dummies 20,000 obs

```{r}
case1 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "s2_lai",
                      subsampling = F, save_results = TRUE)
```

## 2. LiDAR LAI without dummies 20,000 obs
```{r}
case2 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "lidar_lai",
                      subsampling = F, save_results = TRUE)
```
## 3. deltaLAI with dummies 20,000 obs
```{r}
case3 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = F, target_var = "deltaLAI",
                      subsampling = F, save_results = TRUE)
```
## 4. deltaLAI without dummies 20,000 obs
```{r}
a
case4 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = FALSE, target_var = "deltaLAI",
                      subsampling = F, save_results = FALSE)
```

## 1. LiDAR LAI with dummies all obs
```{r}
case1 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = TRUE, target_var = "LiDAR_LAI")
```

## 2. LiDAR LAI without dummies all obs
```{r}
case2 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = FALSE, target_var = "LiDAR_LAI")
```

## 3. deltaLAI with dummies all obs
```{r}
case3 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = TRUE, target_var = "deltaLAI")
```

## 4. deltaLAI without dummies all obs
```{r}
case4 <- run_rf_cases(sites = sites, results_dir = results_dir, output_dir,
                      forest_composition = forest_composition, 
                      use_dummies = FALSE, target_var = "deltaLAI")
```

```{r}
run_rf_cases <- function(training_sites, test_site, results_dir, output_dir, 
                         forest_composition, use_dummies, target_var,
                         subsampling = FALSE, save_results = TRUE) {
  
  # Initialize an empty list to store the data frames
  all_predictors <- list()
  
  # Load data for all sites
  for (site in c(training_sites, test_site)) {
    site_data <- load_metrics_for_site(site, results_dir, forest_composition)  # No dummies here
    if (target_var == "deltaLAI") {
      site_data$deltaLAI <- site_data$lidar_lai - site_data$s2_lai
      site_data$lidar_lai <- NULL  # Remove lidar_lai when deltaLAI is the target
    }
    
    all_predictors[[site]] <- site_data  # Store the data frame in the list
  }
  
  # Combine all the data frames into one
  combined_predictors <- do.call(rbind, all_predictors)
  
  # Create a 'train' column to indicate training vs testing data
  combined_predictors$train <- ifelse(combined_predictors$site %in% training_sites, TRUE, FALSE)
  
  # Create training data
  train_data <- combined_predictors[combined_predictors$train == TRUE, ]
  
  # Create test data
  test_data <- combined_predictors[combined_predictors$train == FALSE, ]
  
  # If using dummies, create dummy variables for the 'site' variable
  if (use_dummies) {
    # Convert the 'site' column to a factor
    train_data$site <- as.factor(train_data$site)
    
    # Create dummy variables for 'site' and remove the original 'site' column
    dummy_vars_train <- model.matrix(~ site - 1, data = train_data)
    colnames(dummy_vars_train) <- gsub("site", "site_", colnames(dummy_vars_train))
    train_data <- cbind(train_data[, !colnames(train_data) %in% "site"], as.data.frame(dummy_vars_train))
    
    # Similarly for test data
    test_data$site <- as.factor(test_data$site)
    dummy_vars_test <- model.matrix(~ site - 1, data = test_data)
    colnames(dummy_vars_test) <- gsub("site", "site_", colnames(dummy_vars_test))
    test_data <- cbind(test_data[, !colnames(test_data) %in% "site"], as.data.frame(dummy_vars_test))
  }
  
  # Shuffle
  if (subsampling){
    train_data <- train_data[sample(nrow(train_data), 20000), ]
  } else {
    train_data <- train_data[sample(nrow(train_data)), ]
  }
  
  # Select the appropriate target variable for the RF formula
  if (target_var == "LiDAR_LAI") {
    formula <- lidar_lai ~ .
  } else if (target_var == "deltaLAI") {
    formula <- deltaLAI ~ .
  }
  
  # Fit the Random Forest model using ranger
  rf_model <- ranger(
    formula = formula,
    data = train_data,
    num.trees = 500,
    mtry = floor(ncol(train_data) / 3),
    importance = "none",
    min.node.size = 5,
    num.threads = 12,
    splitrule = "extratrees",
    num.random.splits = 10,
    sample.fraction = 0.3
  )
  
  # Evaluate the model on the test set
  test_predictions <- predict(rf_model, data = test_data)$predictions
  
  # Calculate evaluation metrics (if needed)
  if (target_var == "LiDAR_LAI") {
    test_actuals <- test_data$lidar_lai
  } else if (target_var == "deltaLAI") {
    test_actuals <- test_data$deltaLAI
  }
  
  # Calculate Mean Squared Error (MSE) for the test set
  test_mse <- mean((test_predictions - test_actuals) ^ 2)
  
  ss_res <- sum((test_actuals - test_predictions) ^ 2)
  ss_tot <- sum((test_actuals - mean(test_actuals)) ^ 2)
  test_r_squared <- 1 - (ss_res / ss_tot)
  
  # Extract model evaluation metrics
  oob_mse <- round(rf_model$prediction.error, 2)
  r_squared <- round(rf_model$r.squared, 2)
  
  # Extract variable importance
  # varImp <- sort(rf_model$variable.importance, decreasing = TRUE)
  
  # Plot and save the variable importance plot
  if (save_results) {
    plot_file <- file.path(output_dir, 
                           paste0(target_var, 
                                  ifelse(use_dummies, 
                                         "_with_dummies", "_no_dummies"), 
                                  "_importance.png"))
    png(plot_file, width = 1920, height = 1080, res = 300)
  }
  
  # barplot(varImp, main = paste("Variable Importance for",
  #                              target_var, 
  #                              ifelse(use_dummies, 
  #                                     "with Dummies", "without Dummies")),
  #         las = 2, cex.names = 0.7)
  
  # Add MSE and R² to the plot as text
  mse_text <- paste("OOB MSE:", oob_mse)
  r2_text <- paste("R²:", r_squared)
  
  # Add the text to the plot
  # text(x = 10, y = max(varImp) - 1, 
  #      labels = paste(mse_text, "\n", r2_text), 
  #      pos = 4, cex = 1.6, col = "blue")
  
  if (save_results) {
    dev.off()  # Close the device if saving
  }
  
  # Save the model's key metrics to a file
  if (save_results) {
    metrics_file <- file.path(output_dir, 
                              paste0(target_var, 
                                     ifelse(use_dummies, 
                                            "_with_dummies", "_no_dummies"), 
                                     "_metrics.txt"))
    writeLines(c(
      paste("OOB MSE:", oob_mse),
      paste("Test MSE:", test_mse),
      paste("R Squared:", r_squared)
    ), con = metrics_file)
  } else {
    # Print the results to the console if not saving
    cat("OOB MSE:", oob_mse, "\n")
    cat("Test MSE:", test_mse, "\n")
    cat("R Squared:", r_squared, "\n")
    cat("Test R Squared:", test_r_squared, "\n")
  }
  
  # Remove the model from memory to free up space
  rm(rf_model)
  gc()  # Force garbage collection
  
  return(list(oob_mse = oob_mse,
              test_mse = test_mse,
              var_explained = r_squared
              # variable_importance = varImp
  )
  )
}

```

```{r}
# Define your sites
sites <- c("Aigoual", "Blois", "Mormal")

# Train using Aigoual and Blois, test on Mormal
results <- run_rf_cases(training_sites = sites[1:2], 
                        test_site = sites[3], 
                        results_dir = results_dir, 
                        output_dir = output_dir, 
                        forest_composition = forest_composition, 
                        use_dummies = FALSE, 
                        target_var = "LiDAR_LAI", 
                        subsampling = FALSE, 
                        save_results = FALSE)

# Check the results
print(results)
```

